Solutions to Integer Programming from Quantum Annealing

### 2020-02-08 paper/main.tex
first draft ish

\subsection{Integer Programming Mapping to QUBO}
\label{sec:results:ilp}
The integer quadratic optimization problem can be mapped to a minimization of the quadratic objective function
\begin{align}
\chi^2 = & \left[\Psi^x_{j^\prime} T^x_{j^\prime i}d_{ij} T^x_{jk}\Psi^x_k + c_i T^x_{ij} \Psi^x_j + p (A_{ij} T^x_{jk} \Psi^x_{k} + T^s_{ij} \Psi^s_j + b_i)^2 \right],\\
\Psi = & \begin{pmatrix}
\Psi^x\\
\Psi^s
\end{pmatrix},
\end{align}
where $p$ is the strength of the penalty factor which must be set large enough such that the constraints are satisfied under minimization. The objective function can be represented as a QUBO Hamiltonian
\begin{align}
E = &
\begin{pmatrix}
\Psi^x & \Psi^s
\end{pmatrix}
\begin{pmatrix}
{T^{x}}^T Q_{00} T^x & {T^{x}}^T Q_{01} T^s\\
{T^{s}}^T Q_{10} T^{x} & {T^{s}}^T Q_{11} T^s
\end{pmatrix}
\begin{pmatrix}
\Psi^x\\ \Psi^s
\end{pmatrix} + b^2,\\
\equiv & \Psi^T Q \Psi + C,
\end{align}
where
\begin{align}
Q_{00} = & d + A^T A + \mathrm{Diag} \left(c + A^T b + b^T A\right),\\
Q_{01} = & Q_{10}^T = A^T,\\
Q_{11} = & \mathrm{Diag}\left(1 + b^T + b\right).
\end{align}
The function $\mathrm{Diag}(x) = \mathds{1}\circ(x\otimes\mathbf{1})$ embeds a vector $x$ into a diagonal matrix, and absorbs the linear contributions of the QUBO into the diagonal elements of the quadratic representation.

\subsection{Application to Minimum Dominating Set}
\label{sec:results:mds}

@@

The solution to the minimum independent set can be expressed as an integer optimization problem given by,
\begin{align}
f(x) = &\min(\sum_{v_i\in V} \psi_{v_i}),\\
\textrm{subject to} \quad & \psi_{v_i} + \sum_{v_j \in \mathit{N}(v_i)} \psi_{v_j} \geq 1,\\
& \psi_{v_i} \in \{0, 1\}.
\end{align}
The problem minimizes the number of vertices in $D$ with a binary variable $\psi$ encoded by a single qubit, subject to the constraint that at least one vertex in $\mathit{N}(v_i)$ is in $D$. We introduce slack variables to encode the inequality constraint such that
\begin{align}
f(x) = &\min(\sum_{v_i\in V} \psi_{v_i}),\\
\textrm{subject to} \quad & \psi_{v_i} + s_{v_i} + J_{v_i v_j} \psi_{v_j} = 1,\\
& \psi_{v_i} \in \{0, 1\},\\
& 0 \leq s_{v_i} \leq |\mathrm{N}(v_i)|\\
& s_{v_i} \in \mathbb{Z},
\end{align}
where the nearest-neighbor sum is expressed by $J$, the adjacency matrix for $G$. The algorithm uses $\overline{\overline{V}}$ qubits to encode $\overline{\overline{V}}$ vertices and $\sum_{v_i \in V} \log_2 \mathit(N(v_i))$ qubits to embed the slack variables. Therefore, the embedding at worse scales with $\overline{\overline{V}} \log_2 |\overline{\overline{V}}|$ qubits for fully connected graphs.

>>>>

\subsection{Integer Programming Mapping to QUBO}
\label{sec:results:ilp}
The integer quadratic optimization problem can be mapped to a minimization of the quadratic objective function
\begin{align}
\chi^2 = & \left[\Psi^x_{j^\prime} T^x_{j^\prime i}d_{ij} T^x_{jk}\Psi^x_k + c_i T^x_{ij} \Psi^x_j + p (A_{ij} T^x_{jk} \Psi^x_{k} + T^s_{ij} \Psi^s_j + b_i)^2 \right],\\
\Psi = & \begin{pmatrix}
\Psi^x\\
\Psi^s
\end{pmatrix},
\end{align}
where $p$ is the strength of the penalty factor which must be set large enough such that the constraints are satisfied under minimization. The objective function can be represented as a QUBO Hamiltonian
\begin{align}
E = &
\begin{pmatrix}
\Psi^x & \Psi^s
\end{pmatrix}
\begin{pmatrix}
{T^{x}}^T Q_{00} T^x & {T^{x}}^T Q_{01} T^s\\
{T^{s}}^T Q_{10} T^{x} & {T^{s}}^T Q_{11} T^s
\end{pmatrix}
\begin{pmatrix}
\Psi^x\\ \Psi^s
\end{pmatrix} + pb^2,\\
\equiv & \Psi^T Q \Psi + C,
\label{eq:matrix_form}
\end{align}
where
\begin{align}
Q_{00} = & d + \mathrm{Diag}(c) + p \left[ A^T A + \mathrm{Diag} \left(A^T b + b^T A\right)\right],\\
Q_{01} = & Q_{10}^T = pA^T,\\
Q_{11} = & p\mathrm{Diag}\left(1 + b^T + b\right).
\end{align}
The function $\mathrm{Diag}(x) = \mathds{1}\circ(x\otimes\mathbf{1})$ embeds a vector $x$ into a diagonal matrix, and absorbs the linear contributions of the QUBO into the diagonal elements of the quadratic representation.

\subsection{Application to the Dominating Set Problem}
\label{sec:results:mds}

@@

\subsubsection{Minimum Dominating Set}
The solution to the minimum independent set can be expressed as an integer optimization problem given by,
\begin{align}
f(x) = &\min(\sum_{v_i\in V} \psi_{v_i}),\\
\textrm{subject to} \quad & \psi_{v_i} + \sum_{v_j \in \mathit{N}(v_i)} \psi_{v_j} \geq 1,\\
& \psi_{v_i} \in \{0, 1\}.
\end{align}
The problem minimizes the number of vertices in $D$ with a binary variable $\psi$ encoded by a single qubit, subject to the constraint that at least one vertex in $\mathit{N}(v_i)$ is in $D$. We introduce slack variables $s_{v_i} = T^s_{v_i v_j} \psi^s_{v_j}$ to encode the inequality constraint such that
\begin{align}
f(x) = &\min(\sum_{v_i\in V} \psi_{v_i}),\\
\textrm{subject to} \quad & \psi_{v_i} + T^s_{v_i v_j} \psi^s_{v_j} + J_{v_i v_j} \psi_{v_j} = 1,\\
& \psi_{v_i} \in \{0, 1\},\\
& 0 \leq s_{v_i} \leq |\mathrm{N}(v_i)|\\
& s_{v_i} \in \mathbb{Z},
\end{align}
where the nearest-neighbor sum is expressed by $J$, the adjacency matrix for $G$. The algorithm uses $\overline{\overline{V}}$ qubits to encode $\overline{\overline{V}}$ vertices and $\sum_{v_i \in V} \log_2 \mathit(N(v_i))$ qubits to embed the slack variables. Therefore, the embedding at worse scales with $\overline{\overline{V}} \log_2 |\overline{\overline{V}}|$ qubits for fully connected graphs.

The target Hamiltonian can be written in the convenient matrix form
\begin{align}
E = &
\begin{pmatrix}
\Psi^x & \Psi^s
\end{pmatrix}
\begin{pmatrix}
1 + p\left[1+J^T J + 2(J+J^T)\right] &p(1+J^T)T^s\\
p{T^s}^T(1+J)& p\left[{T^s}^T T^s + {T^s}^T + T^s\right]
\end{pmatrix}
\begin{pmatrix}
\Psi^x\\ \Psi^s
\end{pmatrix} + p.
\label{eq:matrix_form}
\end{align}

\subsubsection{Total Dominating Set}
